---
title: Ideological Identification
---


*ALL YEARS EXC. 2000 TELEPHONE,2002:* 
We hear a lot of talk these days about liberals and conservatives. Here is (1972,1974: I’m going to show you) a 7-point scale on which the political views that people might hold are arranged from extremely liberal to extremely conservative.

Where would you place yourself on this scale, or haven’t you thought much about this? (7-POINT SCALE SHOWN TO R)

*2000 TELEPHONE:*
When it comes to politics, do you usually think of yourself as extremely liberal, liberal, slightly liberal, moderate or middle of the road, slightly conservative, extremely conservative, or haven’t you thought much about this?

*2002:*
We hear a lot of talk these days about liberals and conservatives. When it comes to politics, do you usually think of yourself as extremely liberal, liberal, slightly liberal, moderate or middle of the road, slightly conservative, extremely conservative, or haven’t you thought much about this?

*2012 NOTE:*
Due to an error in the technical specifications, option ‘Haven’t thought much about it’ was not included in the web instrument; code 9 represents FTF cases only.

## Overall Trend

```{r}
library(tidyverse)
library(ggpubr)
library(plotly)
library(gt)
options(digits=3)

anes_df <- read_csv(here::here("data", "anes_timeseries_cdf_csv_20220916.csv"))

pal <- colorRampPalette(c("orangered3", "gray","darkblue"))
pal <- rev(c(pal(7)[1:3],"springgreen4", pal(7)[4:7]))

p <- anes_df |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    #filter(VCF0127b %in% c(1, 3) & VCF0803 > 0) |>
    group_by(VCF0004, VCF0803) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |> 
    rename("Year"="VCF0004") |>
    ggplot(aes(x=Year, y=Percentage, 
    fill=Party)) + 
    geom_col() + 
    geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
    geom_hline(yintercept=c(25, 75), linetype="dotted", color="grey70") + 
    scale_fill_manual("Party ID", values=pal) + 
    labs(y="Percentage", x="") + 
    theme_pubr() +
    theme(legend.position="bottom")
        
```

:::{.toggle-div .visible}
```{r ideo-all}
p
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h"
        )
    ) 
p
```

:::


## By Education


```{r}

levels <- 1:4
variable <- "VCF0110"
var_levels <- c("Less than High School", "High School Degree", 
        "Some College", "College Degree")
var_name <- "Education"
p <- list()

p_data <- anes_df |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    rename(var=!!variable) |>
    group_by(VCF0004, VCF0803,var) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004, var) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |>
    rename("Year"="VCF0004")

for(ii in seq_along(levels)){
    p[[ii]] <- p_data |> 
        filter(total > 50) |>
        filter(var==ii) |>
        ggplot(aes(x=Year, y=Percentage, 
        fill=Party)) + 
        geom_col() + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_manual("Party ID", values=pal) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")

}

```

::: {.panel-tabset group="education"}

### Less Than High School 

::::{.toggle-div .visible}
```{r ideo-ed-less-than}
p[[1]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[1]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### High School Degree

::::{.toggle-div .visible}
```{r ideo-ed-high-school}
p[[2]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[2]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Some College


::::{.toggle-div .visible}
```{r ideo-ed-some-college}
p[[3]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[3]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### College Degree

::::{.toggle-div .visible}
```{r ideo-ed-college}
p[[4]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[4]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Distribution

```{r}

p <- p_data |> filter(VCF0803==1) |> 
    group_by(Year) |> 
    mutate(Percentage= total/sum(total)*100, 
        !!var_name:=factor(var, labels=var_levels, levels=levels)) |> 
    ggplot(aes(x=Year, y=Percentage, 
        fill=!!sym(var_name))) + 
        geom_col(color="gray", linewidth=.5) + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_viridis_d(option="H", begin=.05, end=.95) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
        theme(legend.position="bottom")
```


::::{.toggle-div .visible}
```{r ideo-ed-dist}
p + guides(fill=guide_legend(nrow=2))
```

::::


::::{.toggle-div }

```{r}

ggplotly(p) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Data

```{r}

data_out <- p_data |> ungroup() |>
    select(Year, var, Party, Percentage, total, n) |> 
    mutate(var=factor(var, levels, 
        labels=var_levels)) |> 
    rename("Group N"=total, "N"=n) |> 
    drop_na(var) |> 
    arrange(Year, var, Party) |> 
    rename(!!var_name:=var) 

data_out |> 
    gt() |> 
    fmt_integer(columns=c("N", "Group N")) |> 
    fmt_number(columns=Percentage) |>
    opt_interactive(use_filters=TRUE, use_compact_mode=TRUE)

write_csv(file=here::here("voter_trends", "data", paste0("ideo_",var_name, ".csv")), data_out)

```

The summary statistics are available [here](data/ideo_Education.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `Party` column.

Note: These are all _weighted_ statistics (using the `VCF0009z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 


:::


## By Gender


```{r}

levels <- 1:2
variable <- "VCF0104"
var_levels <- c("Male", "Female")
var_name <- "Gender"
p <- list()

p_data <- anes_df |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    rename(var=!!variable) |>
    group_by(VCF0004, VCF0803,var) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004, var) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |>
    rename("Year"="VCF0004")

for(ii in seq_along(levels)){
    p[[ii]] <- p_data |> 
        filter(total > 50) |>
        filter(var==ii) |>
        ggplot(aes(x=Year, y=Percentage, 
        fill=Party)) + 
        geom_col() + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_manual("Party ID", values=pal) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")

}

```

::: {.panel-tabset group="gender"}

### Male 
::::{.toggle-div .visible}
```{r ideo-gender-male}
p[[1]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[1]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Female
::::{.toggle-div .visible}
```{r ideo-gender-female}
p[[2]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[2]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Distribution

```{r}

p <- p_data |> filter(VCF0803==1) |> 
    group_by(Year) |> 
    mutate(Percentage= total/sum(total)*100, 
        !!var_name:=factor(var, labels=var_levels, levels=levels)) |> 
    ggplot(aes(x=Year, y=Percentage, 
        fill=!!sym(var_name))) + 
        geom_col(color="gray", linewidth=.5) + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_viridis_d(option="H", begin=.05, end=.95) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")
```


::::{.toggle-div .visible}
```{r ideo-gender-dist}
p + guides(fill=guide_legend(nrow=2))
```

::::


::::{.toggle-div }

```{r}

ggplotly(p) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Data

```{r}


data_out <- p_data |> ungroup() |>
    select(Year, var, Party, Percentage, total, n) |> 
    mutate(var=factor(var, levels, 
        labels=var_levels)) |> 
    rename("Group N"=total, "N"=n) |> 
    drop_na(var) |> 
    arrange(Year, var, Party) |> 
    rename(!!var_name:=var) 

data_out |> 
    gt() |> 
    fmt_integer(columns=c("N", "Group N")) |> 
    fmt_number(columns=Percentage) |>
    opt_interactive(use_filters=TRUE, use_compact_mode=TRUE)

write_csv(file=here::here("voter_trends", "data", paste0("ideo_",var_name, ".csv")), data_out)

```

The summary statistics are available [here](data/ideo_Gender.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `Party` column.

Note: These are all _weighted_ statistics (using the `VCF0009z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 

:::



## By Income


```{r}

levels <- 1:5
variable <- "VCF0114"
var_levels <- c("0-16 Percentile", "17-33 Percentile",
                "34-67 Percentile", "68-95 Percentile",
                "96-100 Percentile")
var_name <- "Income"
p <- list()

p_data <- anes_df |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    rename(var=!!variable) |>
    group_by(VCF0004, VCF0803,var) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004, var) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |>
    rename("Year"="VCF0004")

for(ii in seq_along(levels)){
    p[[ii]] <- p_data |> 
        filter(total > 50) |>
        filter(var==ii) |>
        ggplot(aes(x=Year, y=Percentage, 
        fill=Party)) + 
        geom_col() + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_manual("Party ID", values=pal) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")

}

```

::: {.panel-tabset group="income"}

### 0-16 Percentile

::::{.toggle-div .visible}
```{r ideo-income-1}
p[[1]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[1]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### 17-33 Percentile

::::{.toggle-div .visible}
```{r ideo-income-2}
p[[2]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[2]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### 14-67 Percentile
::::{.toggle-div .visible}
```{r ideo-income-3}
p[[3]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[3]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::
### 67-95 Percentile
::::{.toggle-div .visible}
```{r ideo-income-4}
p[[4]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[4]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### 96-100 Percentile

::::{.toggle-div .visible}
```{r ideo-income-5}
p[[5]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[5]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Distribution

```{r}

p <- p_data |> filter(VCF0803==1) |> 
    group_by(Year) |> 
    mutate(Percentage= total/sum(total)*100, 
        !!var_name:=factor(var, labels=var_levels, levels=levels)) |> 
    ggplot(aes(x=Year, y=Percentage, 
        fill=!!sym(var_name))) + 
        geom_col(color="gray", linewidth=.5) + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_viridis_d(option="H", begin=.05, end=.95) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")
```


::::{.toggle-div .visible}
```{r ideo-income-dist}
p + guides(fill=guide_legend(nrow=2))
```

::::


::::{.toggle-div }

```{r}

ggplotly(p) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::



### Data

```{r}


data_out <- p_data |> ungroup() |>
    select(Year, var, Party, Percentage, total, n) |> 
    mutate(var=factor(var, levels, 
        labels=var_levels)) |> 
    rename("Group N"=total, "N"=n) |> 
    drop_na(var) |> 
    arrange(Year, var, Party) |> 
    rename(!!var_name:=var) 

data_out |> 
    gt() |> 
    fmt_integer(columns=c("N", "Group N")) |> 
    fmt_number(columns=Percentage) |>
    opt_interactive(use_filters=TRUE, use_compact_mode=TRUE)

write_csv(file=here::here("voter_trends", "data", paste0("ideo_",var_name, ".csv")), data_out)

```

The summary statistics are available [here](data/ideo_Income.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `Party` column.

Note: These are all _weighted_ statistics (using the `VCF0009z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 


:::



## By Race

Note: Options for race have varied significantly overtime, and the size of the ANES means that there are very small samples Asian and American Indian. Because of this they are grouped together into "Other, non-Hispanic" which also includes those who identify as multiple racial categories. I also drop any years where there are fewer than 50 individuals in the sample. This is far from ideal. 

```{r}

levels <- 1:4
variable <- "VCF0105b"
var_levels <- c("White, non-Hispanic", "Black, non-Hispanic", 
        "Hispanic", "Other, non-Hispanic")
var_name <- "Race"
p <- list()

p_data <- anes_df |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    rename(var=!!variable) |>
    group_by(VCF0004, VCF0803,var) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004, var) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |>
    rename("Year"="VCF0004")

for(ii in seq_along(levels)){
    p[[ii]] <- p_data |> 
        filter(total > 50) |> 
        filter(var==ii) |>
        ggplot(aes(x=Year, y=Percentage, 
        fill=Party)) + 
        geom_col() + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_manual("Party ID", values=pal) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")

}

```

::: {.panel-tabset group="race"}

### White, non-Hispanic

::::{.toggle-div .visible}
```{r ideo-race-white}
p[[1]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[1]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Black, non-Hispanic

::::{.toggle-div .visible}
```{r ideo-race-black}
p[[2]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[2]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Hispanic

::::{.toggle-div .visible}
```{r ideo-race-hispanic}
p[[3]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[3]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Other, non-Hispanic

::::{.toggle-div .visible}
```{r ideo-race-other}
p[[4]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[4]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Distribution

```{r}

p <- p_data |> filter(VCF0803==1) |> 
    group_by(Year) |> 
    mutate(Percentage= total/sum(total)*100, 
        !!var_name:=factor(var, labels=var_levels, levels=levels)) |> 
    ggplot(aes(x=Year, y=Percentage, 
        fill=!!sym(var_name))) + 
        geom_col(color="gray", linewidth=.5) + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_viridis_d(option="H", begin=.05, end=.95) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")
```


::::{.toggle-div .visible}
```{r ideo-race-dist}
p + guides(fill=guide_legend(nrow=2))
```

::::


::::{.toggle-div }

```{r}

ggplotly(p) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::



### Data

```{r}


data_out <- p_data |> ungroup() |>
    select(Year, var, Party, Percentage, total, n) |> 
    mutate(var=factor(var, levels, 
        labels=var_levels)) |> 
    rename("Group N"=total, "N"=n) |> 
    drop_na(var) |> 
    arrange(Year, var, Party) |> 
    rename(!!var_name:=var) 

data_out |> 
    gt() |> 
    fmt_integer(columns=c("N", "Group N")) |> 
    fmt_number(columns=Percentage) |>
    opt_interactive(use_filters=TRUE, use_compact_mode=TRUE)

write_csv(file=here::here("voter_trends", "data", paste0("ideo_",var_name, ".csv")), data_out)

```

The summary statistics are available [here](data/ideo_Race.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `Party` column.

Note: These are all _weighted_ statistics (using the `VCF0009z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 


:::



## By Subjective Class

Subjective class is what an individual _identifies_ as when asked. The question used by ANES is:

>There’s been some talk these days about different social classes.
Most people say they belong either to the middle class or the working
class. Do you ever think of yourself as belonging in one of these classes?
Which one?  

In some years respondents were asked where they identified within "working" or "middle" but not consistently so that has been dropped from this analysis. 

```{r}


levels <- 1:2
variable <- "VCF0148a"
var_levels <- c("Working Class", "Middle Class")
var_name <- "Class"
p <- list()

p_data <- anes_df |> 
    mutate(VCF0148a = 
        case_match(VCF0148a, 
            1:3 ~ 1, 
            4:6 ~ 2, 
            .default=NA)) |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    rename(var=!!variable) |>
    group_by(VCF0004, VCF0803,var) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004, var) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |>
    rename("Year"="VCF0004")

for(ii in seq_along(levels)){
    p[[ii]] <- p_data |> 
        filter(total > 50) |>
        filter(var==ii) |>
        ggplot(aes(x=Year, y=Percentage, 
        fill=Party)) + 
        geom_col() + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_manual("Party ID", values=pal) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")

}

```

::: {.panel-tabset group="class"}

### Working Class
::::{.toggle-div .visible}
```{r ideo-class-working}
p[[1]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[1]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Middle Class
::::{.toggle-div .visible}
```{r ideo-class-middle}
p[[2]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[2]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::

### Distribution

```{r}

p <- p_data |> filter(VCF0803==1) |> 
    group_by(Year) |> 
    mutate(Percentage= total/sum(total)*100, 
        !!var_name:=factor(var, labels=var_levels, levels=levels)) |> 
    ggplot(aes(x=Year, y=Percentage, 
        fill=!!sym(var_name))) + 
        geom_col(color="gray", linewidth=.5) + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_viridis_d(option="H", begin=.05, end=.95) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")
```


::::{.toggle-div .visible}
```{r ideo-class-dist}
p + guides(fill=guide_legend(nrow=2))
```

::::


::::{.toggle-div }

```{r}

ggplotly(p) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::



### Data

```{r}


data_out <- p_data |> ungroup() |>
    select(Year, var, Party, Percentage, total, n) |> 
    mutate(var=factor(var, levels, 
        labels=var_levels)) |> 
    rename("Group N"=total, "N"=n) |> 
    drop_na(var) |> 
    arrange(Year, var, Party) |> 
    rename(!!var_name:=var) 

data_out |> 
    gt() |> 
    fmt_integer(columns=c("N", "Group N")) |> 
    fmt_number(columns=Percentage) |>
    opt_interactive(use_filters=TRUE, use_compact_mode=TRUE)

write_csv(file=here::here("voter_trends", "data", paste0("ideo_",var_name, ".csv")), data_out)

```

The summary statistics are available [here](data/ideo_Class.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `Party` column.

Note: These are all _weighted_ statistics (using the `VCF0009z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 

:::



## By Union Status


```{r}


levels <- 1:2
variable <- "VCF0127"
var_levels <- c("Union Household", "Non-Union")
var_name <- "Class"
p <- list()

p_data <- anes_df |> 
    filter(VCF0803 %in% c(1:7,9)) |> 
    rename(var=!!variable) |>
    group_by(VCF0004, VCF0803,var) |>  
    summarize(n=sum(VCF0009z)) |> 
    group_by(VCF0004, var) |>
    mutate(total=sum(n), 
        Percentage=n/total*100, 
        Party=factor(VCF0803, c(1:3,4,9,5:7), 
        labels=c("Ext Lib", "Liberal", "Slightly Lib", "Moderate", "Don't Know", "Slightly Con", "Conservative", "Ext Con"))) |>
    rename("Year"="VCF0004")

for(ii in seq_along(levels)){
    p[[ii]] <- p_data |> 
        filter(total > 50) |>
        filter(var==ii) |>
        ggplot(aes(x=Year, y=Percentage, 
        fill=Party)) + 
        geom_col() + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_manual("Party ID", values=pal) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")

}

```

::: {.panel-tabset group="union"}

### Union Household

::::{.toggle-div .visible}
```{r ideo-union-yes}
p[[1]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[1]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::


### Non-Union
::::{.toggle-div .visible}
```{r ideo-union-no}
p[[2]]
```

::::


::::{.toggle-div }

```{r}

ggplotly(p[[2]]) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::



### Distribution

```{r}

p <- p_data |> filter(VCF0803==1) |> 
    group_by(Year) |> 
    mutate(Percentage= total/sum(total)*100, 
        !!var_name:=factor(var, labels=var_levels, levels=levels)) |> 
    ggplot(aes(x=Year, y=Percentage, 
        fill=!!sym(var_name))) + 
        geom_col(color="gray", linewidth=.5) + 
        geom_hline(yintercept=50, linetype="dashed", color="grey70") + 
        geom_hline(yintercept=c(25,75), linetype="dotted",  color="grey70") + 
        scale_fill_viridis_d(option="H", begin=.05, end=.95) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + theme(legend.position="bottom")
```


::::{.toggle-div .visible}
```{r ideo-unions-dist}
p + guides(fill=guide_legend(nrow=2))
```

::::


::::{.toggle-div }

```{r}

ggplotly(p) |> 
    layout(legend = list(
        orientation = "h"
        )
    )

```

::::



### Data

```{r}


data_out <- p_data |> ungroup() |>
    select(Year, var, Party, Percentage, total, n) |> 
    mutate(var=factor(var, levels, 
        labels=var_levels)) |> 
    rename("Group N"=total, "N"=n) |> 
    drop_na(var) |> 
    arrange(Year, var, Party) |> 
    rename(!!var_name:=var) 

data_out |> 
    gt() |> 
    fmt_integer(columns=c("N", "Group N")) |> 
    fmt_number(columns=Percentage) |>
    opt_interactive(use_filters=TRUE, use_compact_mode=TRUE)

write_csv(file=here::here("voter_trends", "data", paste0("ideo_",var_name, ".csv")), data_out)

```

The summary statistics are available [here](data/ideo_Union.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `Party` column.

Note: These are all _weighted_ statistics (using the `VCF0009z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 

:::


## Data Notes

All data presented here is from the [American National Election Studies survey](https://electionstudies.org/) using the `VCF0009z` weights. 
