---
title: School Segregation or Desegregation
---

```{r setup}
library(tidyverse)
library(ggpubr)
library(plotly)
library(gt)
options(digits=3)

anes_df <- read_csv(here::here("data", "anes_timeseries_cdf_csv_20220916.csv"))

pal <- colorRampPalette(c("orangered3", "gray","darkblue"))

source(here::here("data", "scripts.r"))



```


## Question Language

(1964,1968-1972, 1976: What about you?)
Are you in favor of desegregation, strict segregation, or something in between?

1. Desegregation
2. In between
3. Strict segregation
9. Don’t know

## All

```{r}


levels <- c(1:3, 9)
variable <- "VCF0815"
var_levels <- c("Desegregation",
                "In between", "Strict Segregation", 
                "Don't Know")
var_name <- "School Segregation"



all_df <- subgroup_analysis(data=anes_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, middle_value="Don't Know", 
                weight="VCF0011z")
    
all_df <- all_df |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(all_df, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(var_name)))  +
        scale_color_viridis_d(option="H", labels=scales::label_wrap(25)) + 
        # scale_y_continuous(labels=scales::label_percent()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
        theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-all}
p + guides(color=guide_legend(nrow=2)) 

``` 

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    ) 
p

all_df <- all_df |> select(-var) |> 
    rename("Group N"=total, "N"=n ) |> 
    mutate(Group="All") |> 
    ungroup()

```

::: 
## By Education

```{r}
group_var_name <- "Education"
group_var <- "VCF0110"
group_var_levels <- 1:4
group_var_labels <- c("Less than High School", "High School Degree", 
        "Some College", "College Degree")
sub_df <- subgroup_analysis(data=anes_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var=group_var, 
                group_var_name=group_var_name,
                group_var_levels=group_var_levels, 
                group_var_labels=group_var_labels, 
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)

plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-education}
p + guides(color=guide_legend(nrow=2)) 
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    ) 
p

all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

::: 

## By Gender

```{r}



group_var_name <- "Gender"
sub_df <- subgroup_analysis(data=anes_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var="VCF0104", 
                group_var_name=group_var_name,
                group_var_levels=1:2, 
                group_var_labels=c("Male", "Female"),
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)

plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-gender}
p
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    ) 
p
all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

:::


## By Income

```{r}
group_var_name <- "Income"
group_var <- "VCF0114"
group_var_levels <- 1:5
group_var_labels <- c("0-16 Percentile", "17-33 Percentile",
                "34-67 Percentile", "68-95 Percentile",
                "96-100 Percentile")
sub_df <- subgroup_analysis(data=anes_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var=group_var, 
                group_var_name=group_var_name,
                group_var_levels=group_var_levels, 
                group_var_labels=group_var_labels,
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)


plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-income}
p + guides(color=guide_legend(nrow=2)) 
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    )  
p

all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

:::


## By Race

```{r}
group_var_name <- "Race"
group_var <- "VCF0105b"
group_var_levels <- 1:4
group_var_labels <- c("White, non-Hispanic", "Black, non-Hispanic", 
        "Hispanic", "Other, non-Hispanic")
sub_df <- subgroup_analysis(data=anes_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var=group_var, 
                group_var_name=group_var_name,
                group_var_levels=group_var_levels, 
                group_var_labels=group_var_labels,
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)


plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-race}
p
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    ) 
p

all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

:::

Note: Options for race have varied significantly overtime, and the size of the ANES means that there are very small samples Asian and American Indian. Because of this they are grouped together into "Other, non-Hispanic" which also includes those who identify as multiple racial categories. I also drop any years where there are fewer than 50 individuals in the sample. This is far from ideal.


## By Subjective Class

```{r}
group_var_name <- "Class"
group_var <- "VCF0148a"
group_var_levels <- 1:2
group_var_labels <- c("Working Class", "Middle Class")

tmp_df <- anes_df |> 
    mutate(VCF0148a = 
        case_match(VCF0148a, 
            1:3 ~ 1, 
            4:6 ~ 2, 
            .default=NA))

sub_df <- subgroup_analysis(data=tmp_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var=group_var, 
                group_var_name=group_var_name,
                group_var_levels=group_var_levels, 
                group_var_labels=group_var_labels,
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)


plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-class}
p
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    )  
p

all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

:::

Subjective class is what an individual _identifies_ as when asked. The question used by ANES is:

>There’s been some talk these days about different social classes.
Most people say they belong either to the middle class or the working
class. Do you ever think of yourself as belonging in one of these classes?
Which one?  

In some years respondents were asked where they identified within "working" or "middle" but not consistently so that has been dropped from this analysis. 





## By Party

```{r}
group_var_name <- "Partisanship"
group_var <- "VCF0303"
group_var_levels <- 1:3
group_var_labels <- c("Democrats", "Independents", "Republicans")

tmp_df <- anes_df |> 
    mutate(VCF0148a = 
        case_match(VCF0148a, 
            1:3 ~ 1, 
            4:6 ~ 2, 
            .default=NA))

sub_df <- subgroup_analysis(data=tmp_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var=group_var, 
                group_var_name=group_var_name,
                group_var_levels=group_var_levels, 
                group_var_labels=group_var_labels,
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)


plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-party}
p
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    ) 
p

all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

:::

Note: This includes "leaners" in the party they lean towards. Individuals were first asked whether they think of themselves as a Republican, Democrat or Independent. Those who identify as Independent are then asked if they lean towards one party or the other. 

## By Union Status

```{r}
group_var_name <- "Union"
group_var <- "VCF0127"
group_var_levels <- 1:2
group_var_labels <-c("Union Household", "Non-Union")
sub_df <- subgroup_analysis(data=anes_df, 
                variable=variable,
                var_name=var_name,
                var_levels=var_levels, 
                levels=levels, group_var=group_var, 
                group_var_name=group_var_name,
                group_var_levels=group_var_levels, 
                group_var_labels=group_var_labels,
                weight="VCF0011z")

plot_data <- sub_df |> drop_na(!!sym(group_var_name)) |> 
            filter(total > 50)


plot_data <- plot_data |> 
        mutate(`95% Lower`=Percentage - 1.96 * `Standard Error`, 
                 `95% Upper`=Percentage + 1.96 * `Standard Error`)

p <- ggplot(plot_data, aes(x=Year, y=Percentage, 
            ymin=`95% Lower`,
            ymax=`95% Upper`,
        color=!!sym(group_var_name))) + 
        geom_point(position=position_dodge(width=1)) + 
        geom_line(position=position_dodge(width=1)) +
        geom_linerange(position=position_dodge(width=1), 
            alpha=.25,
            aes(group=!!sym(group_var_name)))  +
        scale_color_viridis_d(option="H") + 
        facet_wrap(vars(!!sym(var_name)), labeller=label_wrap_gen()) +
        labs(y="Percentage", x="") + 
        theme_pubr() + 
                theme(legend.position="bottom", 
                legend.justification = "left")
        
```

:::{.toggle-div .visible}
```{r school-segregation-union}
p
```

:::


:::{.toggle-div }

```{r}

p <- ggplotly(p)  |> 
    layout(legend = list(
        orientation = "h",
        y=-0.2
        )
    ) 
p

all_df <- plot_data |> ungroup() |> 
    select(-var, -group) |> 
    rename("Group N"=total, "N"=n, 
            "Group"=!!group_var_name) |> 
    rbind(all_df)
```

:::


## Data 

```{r}

write_csv(file=here::here("voter_trends", "data", paste0("opinion_",var_name, ".csv")), all_df)

```

The summary statistics are available [here](data/opinion_School Segregation.csv). The `Group N` column is the number of respondents in a given group in a given year while the `N` column is the number of respondents who selected the answer in the `School Segregation` column.

Note: These are all _weighted_ statistics (using the `VCF0011z` weights) and are provided if you'd like to make you own visualizations. If you are interested in doing your own analysis, use the raw data available from the [American National Election Studies survey](https://electionstudies.org/). 

